Typing a linear \picalc

### 2020-04-13 paper/paper.tex
Introduce subject congruence

@@

@@

\begin{nitheorem}
  If $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
\end{nitheorem}

@@

\begin{figure}[ht]
  \begin{subfigure}{.5\textwidth}
  \centering
  \begin{tikzpicture}
    \node (gamma) at (0,2) {$\Gamma$};
    \node (psi-m) at (4,2) {$\Psi_m$};
    \node (psi-l) at (4,4) {$\Psi_l$};
    \node (psi-r) at (4,0) {$\Psi_r$};
    \draw[->]                (gamma) -- node[above,sloped] {$\Delta$} node[below] {$\Delta_i = \ell_{\o}$} (psi-m);
    \draw[->]                (gamma) -- node[above,sloped] {$\vdash P$}               (psi-l);
    \draw[->,densely dotted] (gamma) -- node[below,sloped] {$\vdash \subst{P}{j}{i}$} (psi-r);
    \draw[->]                (psi-m) -- node[right]        {$m@i$}                    (psi-l);
    \draw[->]                (psi-m) -- node[right]        {$m@j$}                    (psi-r);
  \end{tikzpicture}
  \caption{Diagrammatic representation of the generalized substitution theorem.}
  \label{fig:subst}
  \end{subfigure}
  % Do not leave empty line
  \begin{subfigure}{.5\textwidth}
  \centering
  \begin{tikzpicture}
    \node (gamma) at (0,2) {$\Gamma$};
    \node (xi-l)  at (0,4) {$\Xi_l$};
    \node (xi-r)  at (0,0) {$\Xi_r$};
    \node (psi-m) at (4,2) {$\Psi_m$};
    \node (psi-l) at (4,4) {$\Psi_l$};
    \node (psi-r) at (4,0) {$\Psi_r$};
    \draw[->]                (gamma) -- node[above,sloped] {$\Delta$} node[below] {$\Delta_i = \ell_{\o}$} (psi-m);
    \draw[->]                (gamma) -- node[left]         {$\ell_i@i$}               (xi-l);
    \draw[->,densely dotted] (gamma) -- node[left]         {$\ell_i@j$}               (xi-r);
    \draw[->]                (xi-l)  -- node[above]        {$\vdash P$}               (psi-l);
    \draw[->,densely dotted] (xi-r)  -- node[above]        {$\vdash \subst{P}{j}{i}$} (psi-r);
    \draw[->]                (psi-m) -- node[right]        {$m@i$}                    (psi-l);
    \draw[->]                (psi-m) -- node[right]        {$m@j$}                    (psi-r);
  \end{tikzpicture}
  \caption{
    Diagrammatic representation of the generalized substitution for process input.
    Thanks to the additional premise in Remark \ref{re:delta-zero}, we know that $m$ is at least capable of $\ell_i$, or $\opsquared{m}{\delta}{\ell_i}$.
  }
  \label{fig:subst-input}
  \end{subfigure}
\end{figure}

>>>>

\todo{lemma for lift-lower}

@@

\begin{lemma}
\todo{introduce entails-unused}
\end{lemma}

@@

Structural congruence is an equivalence relation that aims to overcome the semantically meaningless difference introduced by the syntax of the \picalc{}.

\begin{nitheorem}
  Applying structural congruence rules to a well-typed process should preserve its well-typedness under the same typing and usage contexts.
  If $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
\end{nitheorem}

\begin{proof}
  By induction on \textsc{EQUALS}.
  The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
  For those, we proceed by induction on \textsc{StructCong}.
  The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
  The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (\S \ref{framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
  The middle-point of the resulting derivation on $\comp{Q}{P}$ will no longer be the same.
  The case $\constr{comp-end}$ is trivial because in the typing rule for process $\PO$ the input and output contexts are one and the same.
  In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\ell_{\o}$, making the proof trivial.
  In the opposite direction, we will need to instantiate the newly created channel to a concrete type and multiplicities.
  For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
  In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
  We will then use strengthening (\S \ref{strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
  In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
  We will then use Lemma \ref{lift-lower} to show that lifting a lowered $P$ leaves $P$ unchanged.
  Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in \S \ref{swapping}.
  In the reverse direction, we will use swapping and then show that swapping indices $i$ and $\suc i$ in process $P$ two times leaves $P$ unchanged.
\end{proof}
\todo{make sure constructor names in code coincide}

@@

\begin{figure}[ht]
  \begin{subfigure}{.5\textwidth}
  \centering
  \begin{tikzpicture}
    \node (gamma) at (0,2) {$\Gamma$};
    \node (psi-m) at (4,2) {$\Psi_m$};
    \node (psi-l) at (4,4) {$\Psi_l$};
    \node (psi-r) at (4,0) {$\Psi_r$};
    \draw[->]                (gamma) -- node[above,sloped] {$\Delta$} node[below] {$\Delta_i = \ell_{\o}$} (psi-m);
    \draw[->]                (gamma) -- node[above,sloped] {$\vdash P$}               (psi-l);
    \draw[->,densely dotted] (gamma) -- node[below,sloped] {$\vdash \subst{P}{j}{i}$} (psi-r);
    \draw[->]                (psi-m) -- node[right]        {$m@i$}                    (psi-l);
    \draw[->]                (psi-m) -- node[right]        {$m@j$}                    (psi-r);
  \end{tikzpicture}
  \caption{Diagrammatic representation of the generalized substitution theorem.}
  \label{fig:subst}
  \end{subfigure}
  % Do not leave empty line
  \begin{subfigure}{.5\textwidth}
  \centering
  \begin{tikzpicture}
    \node (gamma) at (0,2) {$\Gamma$};
    \node (xi-l)  at (0,4) {$\Xi_l$};
    \node (xi-r)  at (0,0) {$\Xi_r$};
    \node (psi-m) at (4,2) {$\Psi_m$};
    \node (psi-l) at (4,4) {$\Psi_l$};
    \node (psi-r) at (4,0) {$\Psi_r$};
    \draw[->]                (gamma) -- node[above,sloped] {$\Delta$} node[below] {$\Delta_i = \ell_{\o}$} (psi-m);
    \draw[->]                (gamma) -- node[left]         {$\ell_i@i$}               (xi-l);
    \draw[->,densely dotted] (gamma) -- node[left]         {$\ell_i@j$}               (xi-r);
    \draw[->]                (xi-l)  -- node[above]        {$\vdash P$}               (psi-l);
    \draw[->,densely dotted] (xi-r)  -- node[above]        {$\vdash \subst{P}{j}{i}$} (psi-r);
    \draw[->]                (psi-m) -- node[right]        {$m@i$}                    (psi-l);
    \draw[->]                (psi-m) -- node[right]        {$m@j$}                    (psi-r);
  \end{tikzpicture}
  \caption{
    Diagrammatic representation of the generalized substitution for process input.
    Thanks to the additional premise in Remark \ref{re:delta-zero}, we know that $m$ is at least capable of $\ell_i$, i.e. $\opsquared{m}{\delta}{\ell_i}$ for some $\delta$.
  }
  \label{fig:subst-input}
  \end{subfigure}
\end{figure}
